A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-03-04 sections/conclusion/main.tex
Change the overall structure
New structure:
1. Intro + Motvation + Major cotnributions
2. Lit review + Major contributions
3. Queueing model (with performance measures formulas)
4. Game model (with methodology and numerical results)
5. Conclusion

This study explores a generic 3-player game theoretic model between the 
decision makers of two queueing systems and a service that distributes 
individuals to these two systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model 
that has a tandem buffer and a single service centre (section 
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model 
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time}, 
\ref{sec:proportion_within_target}). 
This novel queuing model is the first contribution of the paper.
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS (section 
\ref{sec:ems_ed_application}).
The inefficiencies that emerge from the perspective of the EMS were explored 
along with ways to apply some incentive mechanisms to improve them.
The key findings from this paper that were observed when playing the game
between two EDs and the EMS (section \ref{sec:application}) are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally
    \item Targeted incentivisation of behaviours can help escape inefficiencies
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed 
that inefficient scenarios can arise by playing the game while the latter 
implies that the learned inefficiencies can be escaped by carefully applying 
certain incentives to the players.
This applied game theoretic model is the second main contribution of this paper.

>>>>

This study explores a generic 3-player game theoretic model between the 
decision makers of two queueing systems and a service that distributes 
individuals to these two systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model 
that has a tandem buffer and a single service centre (section 
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model 
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time}, 
\ref{sec:proportion_within_target}). 
This novel queuing model is the first contribution of the paper.
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS.
The inefficiencies that emerge from the perspective of the EMS were explored 
along with ways to apply some incentive mechanisms to improve them.
The key findings from this paper that were observed when playing the game
between two EDs and the EMS are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally
    \item Targeted incentivisation of behaviours can help escape inefficiencies
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed 
that inefficient scenarios can arise by playing the game while the latter 
implies that the learned inefficiencies can be escaped by carefully applying 
certain incentives to the players.
This applied game theoretic model is the second main contribution of this paper.

### 2022-03-04 sections/introduction/main.tex
Fix some of the `TODOs` (#17)
* Add sections to README.md

* Minor fix in Markov chain formulas

For all formulas consider the arriving state [A_i (u,v)] for individuals
from class i rather than just (u,v). Basically when an individual
arrives at state (u,v) the model will move to a new state based on
that individual's type. This is what this change corresponds to.

* Minor changes on waiting zones

- Display waiting zones 1 and 2 on the queueing system
- Fix some text that listed the incorrect waiting zones

* Changes based on self review

I just read the whole thing and made some minor
chang

% TODO: If this is one paper remove the queueing model part

Emergency departments (EDs) are under increasing pressure to meet patient
waiting time targets and satisfy 
regulations~\cite{EmergencyDepartmentWinterPressures}.
It is widely reported (e.g.~\cite{mirror, thenews, bmj}) that ED congestion 
severely impacts not only patients in the ED but also Emergency Medical 
Services (EMS).
A major concern for ambulances is that they are held waiting parked outside the
ED to offload (dispatch) their patient when the ED is particularly 
busy~\cite{clarey2014ambulance}. 
Since the patient waiting time in ED is measured from the time they enter the
ED itself, there is no incentive, should the patient be stable in the
ambulance, to offload them from EMS to ED services.
As a result, ambulance blocking not only impacts on patients waiting for ED
service, but has a major knock-on effect to delaying the ability of ambulances
to respond to new EMS calls, thus placing lives at risk~\cite{eastanglia}.
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.

A number of papers have been published that touch upon the use of 
queueing models together with game theoretic concepts.
In~\cite{FirmCompetition} the authors study a simultaneous price competition 
between two firms that are modelled as two distinct queueing systems with a 
fixed capacity and a combined arrival rate.
They calculate the Nash equilibrium both for identical and heterogeneous firms
and show that for the former a pure Nash equilibrium always exist and for the 
latter a unique equilibrium exists where only one firm operates.
The authors have also extended their model in~\cite{FirmCompetition2} by 
allowing the players (firms) to choose capacities. 
A main result from this paper was that when both firms operate independently as
a monopoly, the equilibria are socially optimal, but this is not the case when
the firms operate together.
Another extension of~\cite{FirmCompetition} was introduced 
in~\cite{FirmCompetitionExtension} where a long-run version of the competition 
was considered that also had capacity as a decision variable.
In~\cite{knight2017measuring} a normal form game is built that is informed by a 
two-dimensional Markov chain in order to model interactions between critical
care units.
An additional paper that focuses on competition is~\cite{fan2009short} where
the authors created a competition between two sellers where seller 1 supplies 
a product instantly and seller 2 is modelled as a make-to-order M/M/1 queue.
The game that is played requires the two sellers to make a choice on the price 
of the product and then seller 2 to set a capacity that guarantees a maximum 
expected delay.

In the above models, the players are attempting to increase their share of 
individuals choosing to queue.
In public healthcare type settings~\cite{knight2013selfish}, this is not 
necessarily the case. 
Players might not necessarily aim to maximise the so called `market share'.
Rather, the overall service needs to be considered as players aim to minimise
their experienced congestion.
In~\cite{sadat2015can} a healthcare application was studied where patients 
could choose between two hospitals, where a utility function is derived that is
based on patients' perceived quality of life.
How individual behaviour based on congestion impacts the efficiency of public 
services is investigated in~\cite{knight2013selfish}.
The authors place the individuals' choices between different public services 
within the formulation of routing games and measure inefficiencies using a 
concept known as the price of anarchy (PoA)~\cite{koutsoupias1999worst}.
They show that the price of anarchy increases with worth of service and that is
low for systems with insufficient capacities.
In~\cite{deo2011centralized} the authors study the network effect of ambulance 
diversion by proposing a non-cooperative game between two EDs that are modelled
as a queueing network.
Each ED's objective is to minimise its own waiting time and chooses a diversion
threshold based on the patients it has.
In equilibrium both EDs choose to divert ambulances in order to avoid getting
arrivals from the other ED.
In this paper this concept is extended by allowing the ambulance service to 
decide how to distribute its patients among the two EDs.

Another specific part of our research, as described later in the paper, is the 
construction of a queueing system with a tandem buffer and a single service
centre.
In~\cite{d2015pure} the authors explore threshold joining strategies in a 
Markov model that has two tandem queues.
Another example is the one described in~\cite{burnetas2013customer}
where they investigated a network of multiple tandem queues where customers 
decide which queue to attend before joining.
Similarly, in~\cite{bacsar2002stackelberg} the authors examine a network of 
\(N\) tandem M/M/1 queues and with multi-type customers. 
The customers in this paper react to a price \(p\) by picking demand rates that 
maximise utility.
In~\cite{veltman2005equilibrium} a profit maximisation problem is studied that
has two servers; an M/M/1 queue and a parking service providing complementary 
service while the customer is in the first service. 
The providers gain a reward when customers complete both services and no reward 
when they finish one of them.
One of the main conclusions of this study is that by increasing the general 
demand both providers lower their prices to compensate for the increase in wait.
The problem was later extended by~\cite{sun2009equilibrium} where they 
considered arrivals of batches that can share the parking service.
Finally,~\cite{afeche2007decentralized} examines a tandem network of two M/M/1 
queues that are ran by two different profit-maximising service providers.
The network receives three types of customers; those requiring both services, 
customers requiring the first service and customers requiring the second service.
The authors showed that optimal prices also maximise social utility and that
removing two types of customers that don't need both services leads to higher 
profit and lower demand rate.

EDs in the United Kingdom have to follow a set of regulations imposed to 
them by the National Health Service (NHS).
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

@@

This paper first gives an overview of the game theoretic model, then goes on 
to describe a novel queueing model, and then follows with a number of theoretic 
results pertaining to the performance measures of this model. 
These are used to build the utilities of the game.
Finally, there is a section that describes how this game theoretic framework
can be applied at the interface between the EMS and EDs together with some 
analysis on the behaviour that is observed.

>>>>

% Introduction + motivation
Emergency departments (EDs) are under increasing pressure to meet patient
waiting time targets and satisfy 
regulations~\cite{EmergencyDepartmentWinterPressures}.
It is widely reported (e.g.~\cite{mirror, thenews, bmj}) that ED congestion 
severely impacts not only patients in the ED but also Emergency Medical 
Services (EMS).
A major concern for ambulances is that they are held waiting parked outside the
ED to offload (dispatch) their patient when the ED is particularly 
busy~\cite{clarey2014ambulance}. 
Since the patient waiting time in ED is measured from the time they enter the
ED itself, there is no incentive, should the patient be stable in the
ambulance, to offload them from EMS to ED services.
As a result, ambulance blocking not only impacts on patients waiting for ED
service, but has a major knock-on effect to delaying the ability of ambulances
to respond to new EMS calls, thus placing lives at risk~\cite{eastanglia}.
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.

% ED + EMS application
The system defined in section \ref{sec:model_overview} can be applied 
to a healthcare scenario with two such queueing systems.
The two queueing systems can be thought of as two Emergency Departments (ED) 
and the Emergency Medical Services (EMS) as a distributor that distributes 
patients to them.

There are numerous news articles that focus on the problems that arise when 
ambulances stay blocked outside of the hospital for a long amount of 
time~\cite{dailyrecords, staffordshirelive}.
Some news reports also comment on the long idle time of ambulances when 
they are not in use~\cite{herefordtimes} and there are several reports of 
examples where this became an issue for new patients~\cite{southwalesargus}.
A particular article looked into this problem from the point of view of an 
ambulance paramedic:

\begin{quotation}
    % <!--alex ignore he-she her-him-->
    ``He knows if a patient arrives more urgently in need of help, his will be 
    pushed back''~\cite{bbcwales}

    ``We used to bring poorly patients in, and we were out on the road again in
    15 minutes.''~\cite{bbcwales}
\end{quotation}

The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and the patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts patients even if it is relatively 
full, while low threshold means that the ED blocks ambulances more frequently.

% Major contributions
In the United Kingdom the National Heath Services (NHS) sets some regulations 
on ED performance.
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

@@

% Sign posting
This paper first gives an overview of the game theoretic model, then goes on 
to describe a novel queueing model, and then follows with a number of theoretic 
results pertaining to the performance measures of this model. 
These are used to build the utilities of the game.
Finally, there is a section that describes how this game theoretic framework
can be applied at the interface between the EMS and EDs together with some 
analysis on the behaviour that is observed.

### 2022-03-04 sections/model_overview/main.tex
Change the overall structure
New structure:
1. Intro + Motvation + Major cotnributions
2. Lit review + Major contributions
3. Queueing model (with performance measures formulas)
4. Game model (with methodology and numerical results)
5. Conclusion

@@

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[>=stealth, scale=0.8],
        % the rectangle of Queue 1
        \draw (0,0) -- ++(3cm,0) -- ++(0,-1.5cm) -- ++(-3cm,0);
        % The label above Queue 1 -> M
        \node[anchor=north] at (1.5cm, 1cm) {\( 
            \overbrace{\qquad \qquad \qquad \qquad}^{M} 
        \)};
        % The label below Queue 1 -> Parking space
        \node[anchor=north] at (1.5cm, -1.7cm) {\small{Parking space}};

@@

>>>>

The parameters of the model correspond to the following parameters of the ED 
and the EMS:

\begin{itemize}
    \item \( \lambda_2 \): The rate of non-emergency patients that the EMS 
    receives and distributes to EDs
    \item \( \lambda_{1_i} \): The arrival rate of other patients to ED 
    \(i\in\{A, B\}\)
    \item \( \mu_i \): The service rate of patients at ED \(i\in\{A, B\}\)
    \item \( C_i \): The number of available resources (beds) in ED 
    \(i\in\{A, B\}\)  
    \item \( T_i \): The action that ED \(i\in\{A, B\} \) chooses to play which 
    corresponds to the threshold at which they do not accept EMS patients.
    \item \( N_i \): The total patient capacity of ED \(i\in\{A, B\} \)
    \item \( M_i \): The total parking capacity of ED \(i\in\{A, B\} \)
    \item \( R \): The time target for both EDs
    \item \( \alpha \in [0, 1] \) : \textit{Weighting} of blocking time and 
    lost individuals (equation \ref{eq:obj_distributor})
\end{itemize}

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in Section \ref{sec:model_overview}, the interaction between the two EDs 
is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). This corresponds to reacting to experienced 
delays.

@@

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[>=stealth, scale=0.8],
        % the rectangle of Queue 1
        \draw (0,0) -- ++(3cm,0) -- ++(0,-1.5cm) -- ++(-3cm,0);
        % The label above Queue 1 -> M
        \node[anchor=north] at (1.5cm, 1cm) {\( 
            \overbrace{\qquad \qquad \qquad \qquad}^{M} 
        \)};
        % The label below Queue 1 -> Parking space
        \node[anchor=north] at (1.5cm, -1.7cm) {\small{Parking space}};

@@

\input{sections/model_overview/methodology/main.tex}

\input{sections/model_overview/numerical_results/main.tex}
